Low-Thrust Trajectory Design Using Reachability Sets near Asteroid 4769 Castalia

### 2016-08-17 manuscript.tex
Correct some errors in the paper

A popular approach to deal with this divergence is to use a different gravitational model within the circumscribing sphere.
For example, Reference~\citenum{scheeres2000} uses both a polyhedron field and a spherical harmonic expansion to represent the full gravitational field of the body.
The spherical harmonic coefficients are computed from the polyhedron shape model using a constant density assumption~\cite{werner1997}.
This model is applied while close to the body while the harmonic expansion is applied when outside the circumscribing sphere.
Similarly, Reference~\citenum{herrera2014} uses two different harmonic expansion models to represent the gravitational field inside and outside of the circumscribing sphere.
In this case, the coefficients must be carefully chosen to ensure continuity of the gravitational field at the circumscribing sphere.
In both cases, this type of model results in a cumbersome gravity field expression that requires additional constraints to ensure continuity and validity at the radius of the circumscribing sphere.
Any integration software will need to incorporate a switching mechanism between the models when crossing the circumscribing sphere.
Instead, we use a different approach and model the asteroid as a constant-density polyhedron.

In contrast, the polyhedron model results in an exact closed form expression of the gravitational potential field~\cite{werner1994,werner1996}.
This type of model results in the exact potential up to the accuracy of the shape model and a constant density assumption.
However, the calculation of the potential, or the acceleration, requires a summation over every face of the polyhedron. 
As a result, it typically requires a large amount of computation in contrast to the harmonic expansion models. 
However, the formulation is well suited to parallelization and improvements with efficient coding practices. 
Finally, the polyhedron method is well suited to trajectories passing close to the body and offers a simple metric to determine if a particle is inside the asteroid.
The use of the polyhedron model results in a single expression of the gravitational field that is valid everywhere around the body of the asteroid.

% use of low thrust propulsion
The application of optimal control methods for orbital trajectory design is nontrivial.
Frequently, insight into the problem or intuition on the part of the designer is required to determine initial conditions that will converge to the optimal solution.
However, the asteroid system dynamics are nonlinear and exhibit chaotic behaviors.
This makes solving the optimization problem highly dependent on the initial condition.
Similar to the three-body problem, there is an insufficient number of analytical constants to derive an analytical solution in general.
As a result, accurate numerical methods are required to determine optimal solutions.
These methods are critically dependent on an accurate initial guess in order to allow for convergence to an optimal solution.

% Discuss previous work and their drawbacks to determine transfers about asteroids
There has been a large amount of work on the optimal control of spacecraft orbital transfers in the three-body problem~\cite{mingotti2011,grebow2011}.
The dynamics of a spacecraft about small bodies is very similar to that of the three-body problem and much of the theory developed there is also applicable.
Typically, the optimal control problem is solved via direct methods, which approximate the continuous time problem as a parameter optimization problem.
The state and control trajectories are discretly parameterized and solved in the form of a nonlinear optimization problem.
Alternatively, indirect methods apply calculus of variations to derive the necessary conditions for optimality. 
This yields a lower dimensioned problem as compared to the direct approach.
In addition, satisfication of the necessary conditions guarantee local optimality in contrast to directo methods which result in sub-optimal solutions.

% Detail of work - more empasis and add the novelty of our work
The irregular shape of asteroids leads to a richer dynamics as compared to the familiar Kepler's problem.
We model the motion of particles around asteroids using the restricted full two-body problem.
This model has many similarities with the restricted three body problem, and much of the theory developed for the three-body problem is also applicable~\cite{mondelo2010,herrera2014}.

@@

We formulate an optimal control problem to calculate the reachability set on a lower dimensional \Poincare section.
Given an initial condition and fixed time horizon, the reachable set is the set of states attainable, subject to the operational constraints of the spacecraft.
The generation of the reachable set allows for a more systematic method of determining initial conditions and eases the burden on the designer.
The \Poincare section reduces the dimensionality of teh system dynamics to the study of a related discrete update map.
This allows for the design of complex transfer trajectories through the use of a lower dimensional space.
Rather than relying on intuition or insight into the problem, trajectory states are chosen which minimize a distance metric toward a desired target on the \Poincare section.
This simple methodology allows for extended transfer trajectories which iteratively approach a desired target orbit.

In short, the authors present a systematic method of generating optimal transfer orbits about asteroids.
Typically, optimal transfers are generated using a direct optimization method which results in a sub-optimal solution.
This paper present an indirect optimal control formulation to generate the reachability set on a \Poincare section.
Using the reachability set on the \Poincare section allows for a simple method of choosing trajectories which approach a target.
This method allows us to avoid the difficulties inherent in choosing valid initial conditions for the computation of optimal transfer trajectories.
We develop the optimal control formulation and apply this method to a transfer about asteroid 4769 Castalia.

@@

Consider three vectors \( \vecbf{v}_1, \vecbf{v}_2, \vecbf{v}_3 \in \R^{3 \times 1} \), assumed to be ordered in a counterclockwise direction, which define a face.
It is easy to define the three edges of each face as
\begin{align}\label{eq:edges}
    \vecbf{e}_{i+1,i} = \vecbf{v}_{i+1} - \vecbf{v}_i \in \R^{3 \times 1 },
\end{align}
where the index \( i \in \parenth{1,2,3} \) is used to permute all edges of each face.
Since each edge is a member of two faces, there exist two edges which are defined in opposite directions between the same vertices.
We can also define the outward normal vector to face \( f\)  as
\begin{align}\label{eq:face_normal}
    \hat{\vecbf{n}}_f &= \parenth{\vecbf{v}_{2} - \vecbf{v}_1} \times \parenth{\vecbf{v}_{3} - \vecbf{v}_2} \in \R^{3 \times 1},
\end{align}
and the outward facing normal vector to each edge as
\begin{align}\label{eq:edge_normal}
    \hat{\vecbf{n}}_{i+1,i}^f &= \parenth{\vecbf{v}_{i+1} - \vecbf{v}_i} \times \hat{\vecbf{n}}_f .
\end{align}
For each face we define the face dyad \( \vecbf{F}_f \) as
\begin{align}\label{eq:face_dyad}
    \vecbf{F}_f &= \hat{\vecbf{n}}_f \hat{\vecbf{n}}_f \in \R^{3 \times 3}.
\end{align}
Each edge is a member of two faces and has an outward pointing edge normal vector, given in~\cref{eq:edge_normal}, perpendicular to both the edge and the face normal.
For the edge connecting the vectors \( \vecbf{v}_1 \) and \( \vecbf{v}_2 \), which are shared between the faces \(A\) and \( B\), the per edge dyad is given by
\begin{align}\label{eq:edge_dyad}
    \vecbf{E}_{12} = \hat{\vecbf{n}}_A \parenth{\hat{\vecbf{n}}_{12}^A}^T + \hat{\vecbf{n}}_B \hat{\vecbf{n}}_{21}^B \in \R^{3 \times 3}.
\end{align}
The edge dyad \( \vecbf{E}_e  \), is defined for each edge and is a function of the two adjacent faces meeting at that edge.
The face dyad \( \vecbf{F}_f \), is defined for each face and is a function of the face normal vectors.

Let \( \vecbf{r}_i \in \R^{3 \times 1} \) be the vector from the spacecraft to the vertex \( \vecbf{v}_i \) and it's length is given by \( r_i = \norm{\vecbf{r}_i} \in \R^{1} \).
The per-edge factor \( L_e \in \R^{1}\), for the edge connecting vertices \( \vecbf{v}_i \) and \( \vecbf{v}_j \), with a constant length \( e_{ij} \) is
\begin{align}\label{eq:edge_factor}
    L_e &= \ln \frac{r_i + r_j + e_{ij}}{r_i + r_j - e_{ij}}.
\end{align}
For the face defined by the vertices \( \vecbf{v}_i, \vecbf{v}_j, \vecbf{v}_k \) the per-face factor \( \omega_f \in \R^{1} \) is
\begin{align}\label{eq:face_factor}
    \omega_f &= 2 \arctan \frac{\vecbf{r}_i \cdot \vecbf{r}_j \times \vecbf{r}_k}{r_i r_j r_k + r_i \parenth{\vecbf{r}_j \cdot \vecbf{r}_k} + r_j \parenth{\vecbf{r}_k \cdot \vecbf{r}_i} + r_k \parenth{\vecbf{r}_i \cdot \vecbf{r}_j}} .
\end{align}
The gravitational potential due to a constant density polyhedron is given as
\begin{align}\label{eq:potential}
    U(\vecbf{r}) &= \frac{1}{2} G \sigma \sum_{e \in \text{edges}} \vecbf{r}_e \cdot \vecbf{E}_e \cdot \vecbf{r}_e \cdot L_e - \frac{1}{2}G \sigma \sum_{f \in \text{faces}} \vecbf{r}_f \cdot \vecbf{F}_f \cdot \vecbf{r}_f \cdot \omega_f ,
\end{align}
where \( \vecbf{r}_e\) and \(\vecbf{r}_f \) are the vectors from the spacecraft to any point on the respective edge or face, \( G\) is the universal gravitational constant, and \( \sigma \) is the constant density of the asteroid.
Furthermore we can use these definitions to define the attraction, gravity gradient matrix, and Laplacian as
\begin{align}
    \nabla U ( \vecbf{r} ) &= -G \sigma \sum_{e \in \text{edges}} \vecbf{E}_e \cdot \vecbf{r}_e \cdot L_e + G \sigma \sum_{f \in \text{faces}} \vecbf{F}_f \cdot \vecbf{r}_f \cdot \omega_f , \label{eq:attraction}\\
    \nabla \nabla U ( \vecbf{r} ) &= G \sigma \sum_{e \in \text{edges}} \vecbf{E}_e  \cdot L_e - G \sigma \sum_{f \in \text{faces}} \vecbf{F}_f \cdot \omega_f , \label{eq:gradient_matrix}\\
    \nabla^2 U &= -G \sigma \sum_{f \in \text{faces}}  \omega_f \label{eq:laplacian}.
\end{align}

@@

We define a reference frame originating at the center of mass of the asteroid.
The body-fixed reference frame is composed of the unit vectors \( \hat{\vecbf{x}} , \hat{\vecbf{y}}, \hat{\vecbf{z}} \), which are aligned along the principal axes of smallest, intermediate, and largest moment of inertia, respectively.
The body-fixed equations of motion of a massless particle about an arbitrarily rotating asteroid are given by
\begin{align}\label{eq:body_eoms}
    \ddot{\vecbf{r}} + 2 \vecbf{\Omega} \times \dot{\vecbf{r}} + \vecbf{\Omega} \times \parenth{ \vecbf{\Omega} \times \vecbf{r} } + \dot{\vecbf{\Omega}} \times \vecbf{r} = \nabla U(\vecbf{r}) ,
\end{align}
where \( \vecbf{\Omega} \) is the instantaneous angular velocity vector of the asteroid represented in the body-fixed frame, \( \vecbf{r} \) is the position of the particle in the body-fixed frame, and \( U_{\vecbf{r}} \) is the gradient of the gravitational potential~\cite{scheeres2012a}.
We assume that the asteroid rotates at a uniform rate, \( \norm{\vecbf{\Omega}} = \omega \), about the axis of the maximum moment of inertia, \( \vecbf{\Omega} = \omega \hat{\vecbf{z} }\).
As a result, we can represent the equations of motion in scalar form as
\begin{align} \label{eq:eoms}
    \begin{split}
        \ddot{x} - 2 \omega \dot{y} - \omega^2 y &= U_x , \\
        \ddot{y} + 2 \omega \dot{x} - \omega^2 x &= U_y , \\
        \ddot{z} &= U_z .
    \end{split}
\end{align}

In this situation, the state is defined as \( \vecbf{x} = \bracket{\vecbf{r}~\>\vecbf{v}}^T\) with \(\vecbf{r} = \bracket{x~\>y~\>z}^T \in \R^{3\times1}\) and \(\vecbf{v}= \bracket{ \dot{x}~\>\dot{y}~\>\dot{z} }^T \in \R^{3\times1}\) representing the position and velocity with respect to the body-fixed frame, respectively.
We further assume that our spacecraft is capable of exerting a translational acceleration, \( \vecbf{u} \in \R^{3\times1} \), in any direction, while subject to a maximum magnitude constraint, \( \norm{\vecbf{u}} \leq u_m \).
This is typical of many spacecraft which offer full rotational freedom and can direct a potentially varying force or acceleration in any direction.
The equations of motion may be rewritten in state space form as
\begin{align}\label{eq:state_space_eoms}
    \begin{bmatrix} \dot{\vecbf{r}} \\ \dot{\vecbf{v}} \end{bmatrix} &=
    \begin{bmatrix}\vecbf{v} \\ \vecbf{g} \parenth{\vecbf{r}} + \vecbf{h}\parenth{\vecbf{v}} + \vecbf{u} \end{bmatrix} ,
\end{align}
where the terms \(\vecbf{g} \parenth{\vecbf{r}} \) and \( \vecbf{h}\parenth{\vecbf{v}} \) are given by
\begin{align}\label{eq:state_space_terms}
    \vecbf{g}\parenth{\vecbf{r}} = \begin{bmatrix}  U_x + \omega^2 x \\ U_y + \omega^2 y \\ U_z \end{bmatrix} ,\quad
    \vecbf{h}\parenth{\vecbf{r}} = \begin{bmatrix} 2 \omega \dot{y} \\ -2 \omega \dot{x} \\ 0 \end{bmatrix} .
\end{align}

@@

An optimal control problem is defined by the cost function
\begin{align}\label{eq:cost}
    J = -\frac{1}{2} \left( \vecbf{x}(t_f) - \vecbf{x}_{n}(t_f)\right)^T 
    Q
    \left( \vecbf{x}(t_f) - \vecbf{x}_{n}(t_f)\right) ,
\end{align}
where \( \vecbf{x}_n(t_f) \) is the final state of a control-free trajectory, while the term \( \vecbf{x}(t_f)\) is the final state of a trajectory under the influence of the control input.
We use the matrix \( Q = \text{diag} \bracket{1~\>0~\> 1~\> 1~\>0~\>1 } \in \R^{6 \times 6}\) to represent the mapping onto \( \Sigma \).
Maximization of the distance between \( \vecbf{x}_n \) and \(\vecbf{x} \), on the \Poincare section defined in~\cref{eq:poincare_section}, is equivalent to the minimization of \( J \) defined in~\cref{eq:cost}.
We ensure that the trajectories intersect the \Poincare section through the use of terminal constraints.
In addition, we use the terminal constraints to define a specific direction along which we seek to minimize the cost~\cref{eq:cost}.
Since the \Poincare section is four-dimensional, we parameterize a direction in \( \R^4 \)  using three angles \( \phi_1, \phi_2 , \phi_3 \).
The terminal constraints are given in terms of these angles as
\begin{align}\label{eq:terminal_constraints}
    \begin{split}
        m_1 &= y = 0 , \\
        m_2 &= \parenth{\sin \phi_{1_{d}}} \parenth{ x_1^2 + x_2^2 + x_3^2 + x_4^2} - x_1^2 = 0, \\
        m_3 &= \parenth{\sin \phi_{2_{d}}} \parenth{ x_2^2 + x_3^2 + x_4^2} - x_2^2 = 0, \\
        m_4 &= \parenth{\sin \phi_{3_{d}}} \parenth{ 2 x_3^2 + 2 x_3 \sqrt{x_4^2 + 2 x_4^2}} - x_3 - \sqrt{x_4^2 + x_3^2} = 0 ,
    \end{split}
\end{align}
where we make use of the difference states \( \parenth{x_1, x_2 ,x_3, x_4 }\) defined as
\begin{align}\label{eq:diff_states}
    \begin{split}
        x_1 &= x(t_f) - x_n(t_f) , \\
        x_2 &= z(t_f) - z_n(t_f) , \\
        x_3 &= \dot{x}(t_f) - \dot{x}_n(t_f) , \\
        x_4 &= \dot{z}(t_f) - \dot{z}_n(t_f) . \\
    \end{split}
\end{align}
We select the terminal time, \( t_f \), from the time required for the uncontrolled trajectory to return back to the \Poincare section.
The constraint \( m_1 = 0 \) ensures that the terminal state lies on the \Poincare section.
The constraints \( m_2, m_3, m_4 \) are used to define a direction on the \Poincare section.
\Cref{eq:diff_states} represents the difference between our controlled and uncontrolled trajectory on the \Poincare section.
We approximate the entire reachable set by discretization  over the space of angles \(\phi_1, \phi_2, \phi_3 \).
By convention we assume that the angles lie in the following range
\begin{align*}
    \phi_1, \phi_2 &\in [ 0, \pi ) ,\\
    \phi_3 &\in [ 0 , 2 \pi ) ,
\end{align*}
such that we parameterize all directions on the three sphere, \(\S^3\).
Finally, we also incorporate a control acceleration magnitude constraint given by
\begin{align}\label{eq:control_constraint}
    c(\vecbf{u}) = \vecbf{u}^T \vecbf{u} - u_m^2 \leq 0 ,
\end{align}
where \( u_m \) is the maximum acceleration possible by the propulsion system.
This constraint assumes that the control acceleration may be orientated in any direction yet the acceleration magnitude is variable but bounded.
The goal is to determine the control history \( \vecbf{u}(t) \) such that the cost function~\cref{eq:cost} is minimized while subject to the equations of motion~\cref{eq:body_eoms} and the constraints~\cref{eq:control_constraint,eq:terminal_constraints}.

>>>>

A popular approach to deal with this divergence is to use a different gravitational model within the circumscribing sphere.
For example, Reference~\citenum{scheeres2000} uses both a polyhedron field and a spherical harmonic expansion to represent the full gravitational field of the body.
The spherical harmonic coefficients are computed from the polyhedron shape model using a constant density assumption~\cite{werner1997}.
This model is applied while close to the body while the harmonic expansion is applied when outside the circumscribing sphere.
Similarly, Reference~\citenum{herrera2014} uses two different harmonic expansion models to represent the gravitational field inside and outside of the circumscribing sphere.
In this case, the coefficients must be carefully chosen to ensure continuity of the gravitational field at the circumscribing sphere.
In both cases, this type of approach results in a cumbersome gravity field expression that requires additional constraints to ensure continuity and validity at the radius of the circumscribing sphere.
Any integration software will need to incorporate a switching mechanism between the models when crossing the circumscribing sphere.
Instead, we use a different approach and model the asteroid as a constant-density polyhedron.

The polyhedron model results in an exact closed form expression of the gravitational potential field~\cite{werner1994,werner1996}.
This type of model results in the exact potential up to the accuracy of the shape model and a constant density assumption.
However, the calculation of the potential, or the acceleration, requires a summation over every face of the polyhedron. 
As a result, it typically requires a large amount of computation in contrast to the harmonic expansion models. 
However, the formulation is well suited to parallelization and improvements with efficient coding practices. 
Finally, the polyhedron method is well suited to trajectories passing close to the body and offers a simple metric to determine if a particle is inside the asteroid.
The use of the polyhedron model results in a single expression of the gravitational field that is valid everywhere around the body.

% use of low thrust propulsion
The application of optimal control methods for orbital trajectory design is nontrivial.
Frequently, insight into the problem or intuition on the part of the designer is required to determine initial conditions that will converge to the optimal solution.
However, the asteroid system dynamics are nonlinear and exhibit chaotic behaviors.
This makes solving the optimization problem highly dependent on the initial condition.
Similar to the three-body problem, there is an insufficient number of analytical constants to derive an analytical solution in general.
As a result, accurate numerical methods are required to determine optimal solutions.
These methods are critically dependent on an accurate initial guess in order to allow for convergence.

% Discuss previous work and their drawbacks to determine transfers about asteroids
We model the motion of particles around asteroids using the restricted full two-body problem.
The dynamics of a spacecraft about small bodies is very similar to that of the three-body problem.
This model has many similarities with the restricted three body problem, and much of the theory developed for the three-body problem is also applicable~\cite{mondelo2010,herrera2014}.
In addition, there has been a large amount of work on the optimal control of spacecraft orbital transfers in the three-body problem~\cite{mingotti2011,grebow2011}.
Typically, the optimal control problem is solved via direct methods, which approximate the continuous time problem as a parameter optimization problem.
The state and control trajectories are discretly parameterized and solved in the form of a nonlinear optimization problem.
Alternatively, indirect methods apply calculus of variations to derive the necessary conditions for optimality. 
This yields a lower dimensioned problem as compared to the direct approach.
In addition, satisfication of the necessary conditions guarantee local optimality in contrast to directo methods which result in sub-optimal solutions.

@@

We formulate an optimal control problem to calculate the reachability set on a lower dimensional \Poincare section.
Given an initial condition and fixed time horizon, the reachable set is the set of states attainable, subject to the operational constraints of the spacecraft.
The generation of the reachable set allows for a more systematic method of determining initial conditions and eases the burden on the designer.
The \Poincare section reduces the dimensionality of the system dynamics to the study of a related discrete update map.
This allows for the design of complex transfer trajectories on a lower dimensional space.
Rather than relying on intuition or insight into the problem, trajectories are chosen which minimize a distance metric toward a desired target on the \Poincare section.
This simple methodology allows for extended transfer trajectories which iteratively approach a desired target orbit.

In short, the authors present a systematic method of generating optimal transfer orbits about asteroids.
Typically, optimal transfers are generated using a direct optimization method which results in a sub-optimal solution.
This paper present an indirect optimal control formulation to generate the reachability set on a \Poincare section.
Using the reachability set on the \Poincare section allows for a simple method of choosing trajectories which approach a target.
In addition, the reachability set gives an indication of the regions of the phase space accessible to the spacecraft.
This method allows us to avoid the difficulties inherent in choosing valid initial conditions for the computation of optimal transfer trajectories.
We develop the optimal control formulation and apply this method to a transfer about asteroid 4769 Castalia.

@@

Consider three vectors \( \vecbf{v}_1, \vecbf{v}_2, \vecbf{v}_3 \in \R^{3 \times 1} \), assumed to be ordered in a counterclockwise direction, which define a face.
It is easy to define the three edges of each face as
\begin{align}\label{eq:edges}
    \vecbf{e}_{i+1,i} = \vecbf{v}_{i+1} - \vecbf{v}_i \in \R^{3 \times 1 },
\end{align}
where the index \( i \in \parenth{1,2,3} \) is used to permute all edges of each face.
Since each edge is a member of two faces, there exist two edges which are defined in opposite directions between the same vertices.
We can also define the outward normal vector to face \( f\)  as
\begin{align}\label{eq:face_normal}
    \hat{\vecbf{n}}_f &= \parenth{\vecbf{v}_{2} - \vecbf{v}_1} \times \parenth{\vecbf{v}_{3} - \vecbf{v}_2} \in \R^{3 \times 1},
\end{align}
and the outward facing normal vector to each edge as
\begin{align}\label{eq:edge_normal}
    \hat{\vecbf{n}}_{i+1,i}^f &= \parenth{\vecbf{v}_{i+1} - \vecbf{v}_i} \times \hat{\vecbf{n}}_f \in \R^{3 \times 1}.
\end{align}
For each face we define the face dyad \( \vecbf{F}_f \) as
\begin{align}\label{eq:face_dyad}
    \vecbf{F}_f &= \hat{\vecbf{n}}_f \hat{\vecbf{n}}_f \in \R^{3 \times 3}.
\end{align}
Each edge is a member of two faces and has an outward pointing edge normal vector, given in~\cref{eq:edge_normal}, perpendicular to both the edge and the face normal.
For the edge connecting the vectors \( \vecbf{v}_1 \) and \( \vecbf{v}_2 \), which are shared between the faces \(A\) and \( B\), the per edge dyad is given by
\begin{align}\label{eq:edge_dyad}
    \vecbf{E}_{12} = \hat{\vecbf{n}}_A \hat{\vecbf{n}}_{12}^A + \hat{\vecbf{n}}_B \hat{\vecbf{n}}_{21}^B \in \R^{3 \times 3}.
\end{align}
The edge dyad \( \vecbf{E}_e  \), is defined for each edge and is a function of the two adjacent faces meeting at that edge.
The face dyad \( \vecbf{F}_f \), is defined for each face and is a function of the face normal vectors.

Let \( \vecbf{r}_i \in \R^{3 \times 1} \) be the vector from the spacecraft to the vertex \( \vecbf{v}_i \) and it's length is given by \( r_i = \norm{\vecbf{r}_i} \in \R^{1} \).
The per-edge factor \( L_e \in \R^{1}\), for the edge connecting vertices \( \vecbf{v}_i \) and \( \vecbf{v}_j \), with a constant length \( e_{ij} = \norm{\vecbf{e}_{ij}} \in \R^1\) is
\begin{align}\label{eq:edge_factor}
    L_e &= \ln \frac{r_i + r_j + e_{ij}}{r_i + r_j - e_{ij}}.
\end{align}
For the face defined by the vertices \( \vecbf{v}_i, \vecbf{v}_j, \vecbf{v}_k \) the per-face factor \( \omega_f \in \R^{1} \) is
\begin{align}\label{eq:face_factor}
    \omega_f &= 2 \arctan \frac{\vecbf{r}_i \cdot \vecbf{r}_j \times \vecbf{r}_k}{r_i r_j r_k + r_i \parenth{\vecbf{r}_j \cdot \vecbf{r}_k} + r_j \parenth{\vecbf{r}_k \cdot \vecbf{r}_i} + r_k \parenth{\vecbf{r}_i \cdot \vecbf{r}_j}} .
\end{align}
The gravitational potential due to a constant density polyhedron is given as
\begin{align}\label{eq:potential}
    U(\vecbf{r}) &= \frac{1}{2} G \sigma \sum_{e \in \text{edges}} \vecbf{r}_e \cdot \vecbf{E}_e \cdot \vecbf{r}_e \cdot L_e - \frac{1}{2}G \sigma \sum_{f \in \text{faces}} \vecbf{r}_f \cdot \vecbf{F}_f \cdot \vecbf{r}_f \cdot \omega_f \in \R^1,
\end{align}
where \( \vecbf{r}_e\) and \(\vecbf{r}_f \) are the vectors from the spacecraft to any point on the respective edge or face, \( G\) is the universal gravitational constant, and \( \sigma \) is the constant density of the asteroid.
Furthermore we can use these definitions to define the attraction, gravity gradient matrix, and Laplacian as
\begin{align}
    \nabla U ( \vecbf{r} ) &= -G \sigma \sum_{e \in \text{edges}} \vecbf{E}_e \cdot \vecbf{r}_e \cdot L_e + G \sigma \sum_{f \in \text{faces}} \vecbf{F}_f \cdot \vecbf{r}_f \cdot \omega_f \in \R^{3 \times 1} , \label{eq:attraction}\\
    \nabla \nabla U ( \vecbf{r} ) &= G \sigma \sum_{e \in \text{edges}} \vecbf{E}_e  \cdot L_e - G \sigma \sum_{f \in \text{faces}} \vecbf{F}_f \cdot \omega_f \in \R^{3 \times 3}, \label{eq:gradient_matrix}\\
    \nabla^2 U &= -G \sigma \sum_{f \in \text{faces}}  \omega_f \in \R^1 .\label{eq:laplacian}
\end{align}

@@

We define a reference frame originating at the center of mass of the asteroid.
The body-fixed reference frame is composed of the unit vectors \( \hat{\vecbf{x}} , \hat{\vecbf{y}}, \hat{\vecbf{z}} \), which are aligned along the principal axes of smallest, intermediate, and largest moment of inertia, respectively.
The body-fixed equations of motion of a massless particle about an arbitrarily rotating asteroid are given by
\begin{align}\label{eq:body_eoms}
    \ddot{\vecbf{r}} + 2 \vecbf{\Omega} \times \dot{\vecbf{r}} + \vecbf{\Omega} \times \parenth{ \vecbf{\Omega} \times \vecbf{r} } + \dot{\vecbf{\Omega}} \times \vecbf{r} = \nabla U(\vecbf{r}) ,
\end{align}
where \( \vecbf{\Omega} \in \R^{3 \times 1}\) is the instantaneous angular velocity vector of the asteroid represented in the body-fixed frame, \( \vecbf{r} \) is the position of the particle in the body-fixed frame, and \( \nabla U_{\vecbf{r}} \) is the gradient of the gravitational potential~\cite{scheeres2012a}.
We assume that the asteroid rotates at a uniform rate, \( \norm{\vecbf{\Omega}} = \omega \in \R^1 \), about the axis of the maximum moment of inertia, \( \vecbf{\Omega} = \omega \hat{\vecbf{z} }\).
As a result, we can represent the equations of motion in scalar form as
\begin{align} \label{eq:eoms}
    \begin{split}
        \ddot{x} - 2 \omega \dot{y} - \omega^2 y &= U_x , \\
        \ddot{y} + 2 \omega \dot{x} - \omega^2 x &= U_y , \\
        \ddot{z} &= U_z .
    \end{split}
\end{align}

In this situation, the state is defined as \( \vecbf{x} = \bracket{\vecbf{r}~\>\vecbf{v}}^T \in \R^{6 \times 1}\) with \(\vecbf{r} = \bracket{x~\>y~\>z}^T \in \R^{3\times1}\) and \(\vecbf{v}= \bracket{ \dot{x}~\>\dot{y}~\>\dot{z} }^T \in \R^{3\times1}\) representing the position and velocity with respect to the body-fixed frame, respectively.
We further assume that our spacecraft is capable of exerting a translational acceleration, \( \vecbf{u} \in \R^{3\times1} \), in any direction, while subject to a maximum magnitude constraint, \( \norm{\vecbf{u}} \leq u_m \).
This is typical of many spacecraft which offer full rotational freedom and can direct a potentially varying force or acceleration in any direction.
The equations of motion may be rewritten in state space form as
\begin{align}\label{eq:state_space_eoms}
    \begin{bmatrix} \dot{\vecbf{r}} \\ \dot{\vecbf{v}} \end{bmatrix} &=
    \begin{bmatrix}\vecbf{v} \\ \vecbf{g} \parenth{\vecbf{r}} + \vecbf{h}\parenth{\vecbf{v}} + \vecbf{u} \end{bmatrix} ,
\end{align}
where the terms \(\vecbf{g} \parenth{\vecbf{r}} \) and \( \vecbf{h}\parenth{\vecbf{v}} \) are given by
\begin{align}\label{eq:state_space_terms}
    \vecbf{g}\parenth{\vecbf{r}} = \begin{bmatrix}  U_x + \omega^2 x \\ U_y + \omega^2 y \\ U_z \end{bmatrix} ,\quad
    \vecbf{h}\parenth{\vecbf{r}} = \begin{bmatrix} 2 \omega \dot{y} \\ -2 \omega \dot{x} \\ 0 \end{bmatrix} .
\end{align}

@@

An optimal control problem is defined by the cost function
\begin{align}\label{eq:cost}
    J = -\frac{1}{2} \left( \vecbf{x}(t_f) - \vecbf{x}_{n}(t_f)\right)^T 
    Q
    \left( \vecbf{x}(t_f) - \vecbf{x}_{n}(t_f)\right) ,
\end{align}
where \( \vecbf{x}_n(t_f) \) is the final state of a control-free trajectory, while the term \( \vecbf{x}(t_f)\) is the final state of a trajectory under the influence of the control input.
We use the matrix \( Q = \text{diag} \bracket{1~\>0~\> 1~\> 1~\>0~\>1 } \in \R^{6 \times 6}\) to represent the mapping onto \( \Sigma \).
Maximization of the distance between \( \vecbf{x}_n \) and \(\vecbf{x} \), on the \Poincare section defined in~\cref{eq:poincare_section}, is equivalent to the minimization of \( J \) defined in~\cref{eq:cost}.
We ensure that the trajectories intersect the \Poincare section through the use of terminal constraints.
In addition, we use the terminal constraints to define a specific direction along which we seek to minimize the cost~\cref{eq:cost}.
Since the \Poincare section is four-dimensional, we parameterize a direction in \( \R^4 \)  using three angles \( \phi_1, \phi_2 , \phi_3 \).
The terminal constraints are given in terms of these angles as
\begin{align}\label{eq:terminal_constraints}
    \begin{split}
        m_1 &= y = 0 , \\
        m_2 &= \parenth{\sin \phi_{1_{d}}} \parenth{ x_1^2 + x_2^2 + x_3^2 + x_4^2} - x_1^2 = 0, \\
        m_3 &= \parenth{\sin \phi_{2_{d}}} \parenth{ x_2^2 + x_3^2 + x_4^2} - x_2^2 = 0, \\
        m_4 &= \parenth{\sin \phi_{3_{d}}} \parenth{ 2 x_3^2 + 2 x_3 \sqrt{x_4^2 + 2 x_4^2}} - x_3 - \sqrt{x_4^2 + x_3^2} = 0 ,
    \end{split}
\end{align}
where we make use of the difference states \( \parenth{x_1, x_2 ,x_3, x_4 }\) defined as
\begin{align}\label{eq:diff_states}
    \begin{split}
        x_1 &= x(t_f) - x_n(t_f) , \\
        x_2 &= z(t_f) - z_n(t_f) , \\
        x_3 &= \dot{x}(t_f) - \dot{x}_n(t_f) , \\
        x_4 &= \dot{z}(t_f) - \dot{z}_n(t_f) . \\
    \end{split}
\end{align}
We select the terminal time, \( t_f \), from the time required for the uncontrolled trajectory to return back to the \Poincare section.
The constraint \( m_1 = 0 \) ensures that the terminal state lies on the \Poincare section.
The constraints \( m_2, m_3, m_4 \) are used to define a direction on the \Poincare section.
\Cref{eq:diff_states} represents the difference between our controlled and uncontrolled trajectory on the \Poincare section.
We approximate the entire reachable set by discretization  over the space of angles \(\phi_1, \phi_2, \phi_3 \).
By convention we assume that the angles lie in the following range
\begin{align*}
    \phi_1, \phi_2 &\in [ 0, \pi ) ,\\
    \phi_3 &\in [ 0 , 2 \pi ) ,
\end{align*}
such that we parameterize all directions on the three sphere, \(\S^3\).
Finally, we also incorporate the control acceleration magnitude constraint as
\begin{align}\label{eq:control_constraint}
    c(\vecbf{u}) = \vecbf{u}^T \vecbf{u} - u_m^2 \leq 0 ,
\end{align}
where \( u_m \) is the maximum acceleration possible by the propulsion system.
This constraint assumes that the control acceleration may be orientated in any direction yet the acceleration magnitude is variable but bounded.
The goal is to determine the control history \( \vecbf{u}(t) \) such that the cost function~\cref{eq:cost} is minimized while subject to the equations of motion~\cref{eq:body_eoms} and the constraints~\cref{eq:control_constraint,eq:terminal_constraints}.
